Systematic Specification for Scalable Verification of Fault-Tolerant Algorithms

### 2016-10-20 doc/om-paper/paper.tex
paper edits, tweak SAL file for paper

@@

\lee{finish intro...  One more contribution is that we abstract faults using partially interpreted functions in model-checking, which is new? Note focus is on infinite-state model-checking}

@@

\subsection{Calendar Automata}\label{sec:calendar}\label{sec:calendar}
\emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{}. Timed automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks} in a solver. Timed automata model-checkers are specialized for verifying real-time systems in which the only infinite-valued variables are the real-time clocks. Modern general-purpose model-checkers like \lee{name some} use SMT solvers for general infinite-state system verification that is not limited to real-time clocks.

@@

\subsection{OM(1) Overview}

@@

\begin{figure}
  \centering
  \includegraphics[width=8cm]{figs/proof-structure-validity}
  \caption{Proof structure for the validity property.}
  \label{fig:compare}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=8cm]{figs/proof-structure-agreement}
  \caption{Proof structure for the agreement property.}
  \label{fig:compare}
\end{figure}

@@

>>>>

\usepackage{tikz}
\newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}}

@@

\lee{finish intro...  One more contribution is that we abstract faults using partially interpreted functions in model-checking, which is new? Note focus is on infinite-state model-checking: will need user-provided lemmas. compositional approach to providing lemmas }

@@

\subsection{Calendar Automata}\label{sec:calendar}
\emph{Timed automata} are specialized formalisms for verifying real-time systems~\cite{}. Timed automata assume the existence of continuously-varying real-valued variables, known as \emph{real-time clocks} in a solver. Timed automata model-checkers are specialized for verifying real-time systems in which the only infinite-valued variables are the real-time clocks. Modern general-purpose model-checkers like \lee{name some} use SMT solvers for general infinite-state system verification that is not limited to real-time clocks.

@@

\subsection{OM(1) Overview}
Our first case is the ``Oral Messages'' algorithm to solve the Byzantine
Generals Problem~\cite{Lamport-OM}. The algorithm, $\OM{m}$, is parameterized by
the number of rounds of message relay. Here, we focus on Oral Messages with one
round of communication, $\OM{1}$, that tolerates a single Byzantine fault, so
long as there are a sufficient number of nodes nodes.

Figure~\ref{fig:om1} depicts the communication pattern for \OM{1}. Each of the
$n$ nodes in the system represents a (Byzantine) General. In this formulation, the
center node is the commanding general, while the $n-1$ outer generals are the
lieutenants. The algorithm starts with the commanding general sending each
lieutenant the same message $v$. Upon receipt of the commander's message, each
lieutenant in turn sends the message received to each of the other lieutenants.
Once a lieutenant has received all $n-1$ expected messages, a majority vote is
taken among the $n$ messages and the lieutenant declares its output (course of
action) to be whichever message is in the majority.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
ell/.style={x radius=1, y radius=0.5, draw, shape=circle},
arr/.style={semithick}]
\node[ell] (c)  at (0,0)   {Commander};
\node[ell] (l1) at (0:4)   {Lieutenant 1};
\node[ell] (l2) at (120:4) {Lieutenant 2};
\node[ell] (l3) [fill=gray!30] at (240:4) {Lieutenant 3};
%
\draw[->, arr] (c) -- node[auto] {$v$} (l1);
\draw[->, arr] (c) -- node[auto] {$v$} (l2);
\draw[->, arr] (c) -- node[auto] {$v$} (l3);
\draw[<->, arr] (l1) -- node[auto,swap,near start] {$v$}
                        node[auto,swap,near end] {$v$} (l2);
\draw[<->, arr] (l1) -- node[auto,near start,color=red] {$x$}
                        node[auto,near end] {$v$} (l3);
\draw[<->, arr] (l2) -- node[auto,swap,near start,color=red] {$x$}
                        node[auto,swap,near end] {$v$} (l3);
\end{tikzpicture}
\caption{\OM{1} Four generals, one traitor}
\label{fig:om1}
\end{figure}

For a fault model we follow~\cite{Rushby-OM1} and assume that nodes are the
only source of faults. Moreover, node faults are assumed to be either
\emph{manifest}, \emph{symmetric}, or \emph{byzantine}. A fault is called manifest if
it is detectable by the non-faulty nodes; for example a node that sends to
another node a message that explicitly indicates there is a fault. On the
other hand, a fault is called symmetric if its presence is not necessarily
detectable as a fault to the other components; for example a node which sends
wrong messages, as opposed to invalid or missing messages.

Given a formal model of \OM{1}, the standard target for verification is
\emph{validity} and \emph{agreement}. If we let $l_i$ denote the output for
lieutenant $i$ (where $1 \le i \le n-1$), then validity states:
%
\begin{equation}
    \forall \,i. \quad l_i = v
\end{equation}
%
whereas agreement states:
%
\begin{equation}
    \forall \,i, j. \quad l_i = l_j
\end{equation}
%
where $i,j$ range over the non-faulty nodes.

It is a classical result that \OM{1} can tolerate at most 1 (byzantine) faulty
node as long as there are at least 4 nodes in total. By adding additional
communication rounds, \OM{1} can be extended to an algorithm \OM{m} which
tolerates at most $m$ faults. On the other hand it is known that any system
which tolerates $m$ byzantine faults must involve at least $3m+1$ nodes.

In the results above there are several assumptions made regarding the
underlying computational platform which aren't obvious from the informal
description, but become so when one considers formally modelling it. To make
as many assumptions as possible explicit, we require:
\begin{enumerate}
    \item every message sent by a node is received successfully by the
        addressed node,
    \item when a node receives a message it may determine who sent it,
    \item each node can detect the absence of a message.
\end{enumerate}
These are the assumptions made by Lamport in his constructed solution for
$\OM{m}$ and the corresponding impossibility result (see
\cite{Lamport-OM}~\S2). It follows that a system implementing $\OM{1}$ must
transition synchronously. In particular, in an asynchronous system there is no
general method for detecting when a message is absent.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/proof-structure}
  \caption{Proof structure.}
  \label{fig:proof}
\end{figure}

The proof structure is shown in Figure~\ref{fig:proof}. The two graphs show the structure to prove validity and agreement, respectively. The nodes group related lemmas, and a directed edge from node $a$ to $b$ means lemmas of $b$ depend on $a$. For readability, we have replicated lemmas shared in the proofs of validity and agreement. Additionally, we have elided the dependencies between the lemmas within each node. The proofs are all $k=1$ inductive.

The lemmas are classified into five groups: calendar automata lemmas, abstract state machine lemmas, protocol behavior lemmas, majority vote lemmas, and fault model lemmas.

The calendar automata lemmas do not concern the system. The \emph{cal\_invt} lemma is a conjunction of basic properties about the calendar automata, essentially characterizing the axiomatization described in Section~\ref{sec:calendar}. Additionally, the \emph{missing\_buffer\_size} lemma relates the calendar to a node's message buffer: the messages received added to the empty messages equals the total number of possible messages to receive.

The \emph{abstract\_all} lemmas characterizes the transition between states of the abstract state machine. Technically, \emph{abstract\_all} also depends on ``boilerplate'' lemmas relating each state of the abstract state machine to a state variable representing it, which are elided for space.

The system itself includes both the lemmas characterizing the global protocol behavior as well as majority vote\lee{have we talked about the voting algorithm earlier?}. The global behavior concerns the entire distributed system, whereas the lemmas regarding the majority vote algorithm only concern the behavior of a single node. The proofs mechanize closely the proofs found in Boyer and Moore's original treatment~\cite{}.

Finally, the fault model lemmas relate the fault model to the system. The \emph{fault\_propagation} lemma connects the presence of faulty messages in the calendar with the fault status of nodes that receive those messages in subsequent rounds. The \emph{symmetric\_relay} and \emph{manifest\_relay} lemmas characterize the effect of symmetric and manifest faults, respectively, on the kinds of messages a receiver obtains from a faulty node. All the fault model lemmas mentioned so far are necessary for the agreement proof, but for validity, we need only \emph{candidate\_good}, proving that if the maximum fault assumption holds and the sender is nonfaulty, then no modified messages are received.

@@

% ------------------------------------------------------------
\section{Related Work}\label{sec:related}
